---
title: "Reproducible Research > Week 2 > Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)[52KB] 

The variables included in this dataset are:

 **steps**: Number of steps taking in a 5-minute interval (missing values are coded as `NA`)

 **date**: The date on which the measurement was taken in YYYY-MM-DD format

 **interval**: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Procedure

1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## Code

### 1. Code for reading in the dataset and/or processing the data

#### read data, create baseline data:

```{r readData}
library(scales)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(lubridate)

url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("activity.zip")) download.file(url,"activity.zip",method="curl")
data <- {unzip("activity.zip")}
```

#### transform **interval** into hour + minutes(*2digits* + *2ditits*)
#### create datetime as POSIXct and then create act_df

```{r create df}
read_csv(data,col_types="ncn") %>%
  mutate(interval=formatC(interval,width=4,flag="0")) %>%
  mutate(datetime=as.POSIXct(paste(date,interval),format="%Y-%m-%d %H%M")) %>%
  mutate(date=as.Date(date,"%Y-%m-%d")) -> activity_df
```
```{r}
str(activity_df)
```

### 2. Histogram of the total number of steps taken each day

#### create dataframe:

```{r hist}
activity_df %>% 
  group_by(date) %>% 
  summarise(total= sum(steps,na.rm=TRUE)) %>%
  ungroup(date)-> hist_df
```
```{r}
hist_df
```

#### plot:

```{r plot_hist}
ggplot(hist_df, aes(x=date,y=total)) +
  geom_col(fill="red") +
  ggtitle("Histogram of the total number of steps taken each day") +
             xlab("Date") + ylab("Total Steps")
```

### 3. Mean and median number of steps taken each day

#### create dataframe:

*Median did not appear as it was, so 0 was excluded*

```{r mean and median}
activity_df %>% group_by(date) %>%
  filter(steps!=0) %>%
  summarise(median=median(steps,na.rm=TRUE)) -> median_df

activity_df %>% group_by(date) %>%
  summarise(mean=mean(steps,na.rm=TRUE)) -> mean_df

mean_median_df <- inner_join(mean_df,median_df,by="date") %>%
  gather("summary","count",2:3)
```
```{r}
mean_median_df
```

#### plot:

```{r plot mean_median}
ggplot(mean_median_df) + 
  geom_line(aes(x=date,y=count,color=summary)) +
  ggtitle("Mean and median number of steps taken each day") +
  xlab("Date") + ylab("Steps")
```

### 4. Time series plot of the average number of steps taken

#### create dataframe:
*My time locale is bad, so change to UTC*
```{r average steps}
activity_df %>% group_by(interval) %>% 
  summarise(average=mean(steps,na.rm=TRUE)) %>%
  mutate(interval=as.POSIXct(interval,tz="UTC",format="%H%M"))  -> avg_df
```
```{r}
avg_df
```

#### plot:

```{r plot avg and steps}
ggplot(avg_df) + 
  geom_line(aes(x=interval,y=average),color="red") +
  scale_x_datetime(breaks = date_breaks("4 hour"), labels = date_format("%H:%M")) +
  ggtitle("Time series plot of the average number of steps taken") +
  xlab("Interval") + ylab("Average number of steps") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### The 5-minute interval that, on average, contains the maximum number of steps

#### create dataframe:

```{r avg and max steps}
activity_df %>% group_by(interval) %>% 
  summarise(average=mean(steps,na.rm=TRUE),max=max(steps,na.rm=TRUE)) %>%
  mutate(interval=as.POSIXct(interval,tz="UTC",format="%H%M")) %>%
  ungroup(interval) %>%
  select(interval,average,max) %>%
  gather("summary","count",2:3) -> avg_max_df
```
```{r}
avg_max_df
```

#### plot:

```{r plot avg and max steps}
ggplot() + 
  geom_line(data=avg_max_df,aes(x=interval,y=count,color=summary))  +
  scale_x_datetime(breaks = date_breaks("4 hour"), labels = date_format("%H:%M")) +
  ggtitle("The 5-minute interval that, on average, contains the maximum number of steps") +
  xlab("Interval") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 6. Code to describe and show a strategy for imputing missing data

#### show missing data

```{r missing data}
activity_df %>% filter(is.na(steps)) %>%
  group_by(date) %>%
  summarise(count=n()) %>%
  ungroup(date) 
```

#### show a strategy for imputing missing data
*Assign 0 to NA*
```{r assign 0 to na}
activity_df %>% mutate(steps=if_else(is.na(steps),0,steps))
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed

#### create dataframe:

```{r hist 2}
activity_df %>% 
  mutate(steps=if_else(is.na(steps),0,steps)) %>%
  group_by(date) %>% 
  summarise(total= sum(steps)) %>%
  ungroup(date)-> hist2_df
```
```{r}
hist2_df
```

#### plot:

```{r plot_hist 2}
ggplot(hist2_df, aes(x=date,y=total)) +
  geom_col(fill="blue") +
  ggtitle("Histogram of the total number of steps taken each day 
          after missing values are imputed") +
             xlab("Date") + ylab("Total Steps")
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

### create dataframe:

```{r weekdays and weekends}
activity_df %>%
  mutate(weekday=wday(date,locale="UTC")) %>%
  mutate(week=if_else(weekday %in% c(1:5),"weekday","weekend")) %>%
  group_by(interval,week) %>% 
  summarise(average=mean(steps,na.rm=TRUE)) %>%
  ungroup(interval) %>%
  mutate(interval=as.POSIXct(interval,tz="UTC",format="%H%M")) %>%
  select(interval,week,average) -> week_df
```
```{r}
week_df
```

#### plot:

```{r plot average by weekend}
ggplot(week_df) +
  geom_line(aes(x=interval,y=average,color=week)) +
  scale_x_datetime(breaks = date_breaks("4 hour"), labels = date_format("%H:%M")) +
  ggtitle("Panel plot comparing the average number of steps taken 
          per 5-minute interval across weekdays and weekends") +
  xlab("Interval") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

#### all done